---
title: EWAS Catalog data checks
author: Thomas Battram
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: united
    number_sections: false
    code_folding: hide
params:
  efos: ""
  ids: ""
  dates: ""
  criteria: ""
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}

## pkgs
suppressWarnings(suppressPackageStartupMessages({
library(bookdown) # extra markdown options
library(knitr) # ...
library(kableExtra) # extra nice tables
library(tidyverse) # tidy code and data
library(RColorBrewer) # for plotting colours
library(usefunc) # own package of useful functions
}))

knitr::opts_chunk$set(echo = FALSE, 
                  warning = FALSE, 
                  message = TRUE, 
                  cache = FALSE, 
                  dpi = 300 
                  )
```


```{r general-setup}
# params <- list(efos = "data/bad-efos.RData", ids = "data/bad-ids.RData", dates = "data/bad-dates.RData", criteria = "data/inclusion-criteria-checks.RData")
for (i in names(params)) {
    temp <- new_load(params[[i]])
    assign(i, temp)
}

make_table <- function(tab, columns = NA, cap = NULL, footy = NA)
{
    if (!is.na(columns)) tab <- tab %>% dplyr::select(one_of(columns))
    tidy_tab <- tidy_colnames(tab)
    kab <- kbl(tidy_tab, caption = cap) %>%
        kable_styling()
    if (!is.na(footy)) kab <- kab %>% footnote(general = footy, general_title = "", footnote_as_chunk = T)
    if (nrow(tidy_tab) > 10) kab <- kab %>% scroll_box(width = "100%", height = "450px")
    return(kab)
}

```

## Intro

This report details some checks of the data within The EWAS Catalog. The most serious issues that need to be fixed asap if they occur are:

1. The results data can't be correct - i.e. P value or SE is negative or CpG name doesn't begin with a 'c'
2. The data greatly violates the inclusion criteria - e.g. N = 10 or P = 0.5  
3. If there are duplicated unique identifiers (StudyIDs). 

## Overall checks

```{r overall-checks}
out <- tibble(dup = nrow(ids$dup_studies), 
              pmids = nrow(ids$no_pmid_studies),
              efos = nrow(efos$bad_efo_studies),  
              dates = nrow(dates$bad_date_studies), 
              high_p_studies = criteria$n_high_p_studies, 
              high_p_results = criteria$n_high_p_results,
              low_n = nrow(criteria$low_n_studies), 
              cpg_name = length(unique(criteria$bad_cpg_ids)), 
              neg_p = length(unique(criteria$neg_p_ids)),
              neg_se = length(unique(criteria$neg_se_ids))
              )
make_table(out, cap = "Overall checks", footy = "dup = number of duplcated study IDs, 
  pmids = number of studies without pmids, 
  efos = number of studies with potentially false EFO terms, 
  dates = number of studies with the wrong date format, 
  high-p-studies = number of studies with any results P > 1x10-4, 
  high-p-results = number of results with P > 1x10-4, 
  low-n = number of studies with N < 100, 
  cpg-name = number of studies with results that have CpG names that don't start with a 'c', 
  neg-p = number of studies with negative P values in results, 
  neg-se = number of studies with negative SE in results")
```

## IDs

```{r duplicated-studies}
if (nrow(ids$dup_studies) == 0) {
    message("There are no duplicated studies - woop woop!")
} else {
    make_table(ids$dup_studies, 
           columns = c("Author", "PMID", "Trait", "Analysis", "StudyID"), 
           cap = "Duplicated studies")    
}
```

```{r pmid-studies}
if (nrow(ids$no_pmid_studies) == 0) {
    message("Apart from the EWAS in ARIES, there are no studies without PMIDs - woop woop!")    
} else {
    make_table(ids$no_pmid_studies, 
               columns = c("Author", "PMID", "Trait", "Analysis", "StudyID"), 
               cap = "Studies without a PMID that aren't in-house EWAS")  
}
```

## EFOs

The EFOs that might not exist are: __`r unique(efos$bad_efos)`__. Double check this at: https://www.ebi.ac.uk/efo/. The studies linked to these are below

```{r efo-studies}
if (nrow(efos$bad_efo_studies) == 0) {
    message("There are no studies with EFOs that are not present in the EBI database - woop woop!")    
} else {
    make_table(efos$bad_efo_studies, 
               columns = c("Author", "PMID", "Trait", "EFO", "StudyID"), 
               cap = "Studies with EFOs that were not in the EFO database")  
}
```

## Dates

```{r date-studies}
if (nrow(dates$bad_date_studies) == 0) {
    message("All studies have dates that match the yyyy-mm-dd format - woop woop!")    
} else {
    make_table(dates$bad_date_studies, 
               columns = c("Author", "PMID", "Date", "StudyID"), 
               cap = "Studies with dates are not in the format yyyy-mm-dd")  
}
```

<br>

Any dates that are missing need to be filled in manually. If there are faults in the format of the date you can use the example code below to change the dates:

```{r change-dates-example, eval = FALSE, echo = TRUE}

library(tidyverse)
library(lubridate)

## get data
studies_file <- "" # enter studies file location here
studies <- read_tsv(studies_file, guess_max=1e6)

n_missing <- sum(is.na(studies$Date))

## Code below works if dates are in format dd/mm/yyyy - change accordingly 
yearmd <- ymd(studies$Date)
daymy <- dmy(studies$Date) 
yearmd[is.na(yearmd)] <- daymy[is.na(yearmd)]

## Update the studies file
studies$Date <- yearmd

## Check still same number of missing vals
stopifnot(sum(is.na(studies$Date)) == n_missing)

## Write out the studies file
outfile <- "" # enter output file location here
write.table(studies, file = outfile, row.names = F, col.names = T, quote = F, sep = "\t")

```

## Inclusion criteria

The number of studies and results with at least one high P value (P > 1x10^-4^) and the number of studies with a low N (N < 100) can be seen in __Table \@ref(tab:overall-checks)__. The studies with a low N can be seen below.

```{r lown-studies}
if (nrow(criteria$low_n_studies) == 0) {
    message("No studies have N < 100 - woop woop!")    
} else {
    out <- criteria$low_n_studies %>% arrange(N)
    make_table(out, 
               columns = c("Author", "PMID", "Trait", "Analysis", "N", "StudyID"), 
               cap = "Studies with N < 100")  
}
```

<br>

The number of full studies that would be removed if we removed all results with P > 1x10^-4^ = `r criteria$n_rm_high_p_studies`.

