# Initial EWAS report
### `r Sys.Date()`

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
            warning = FALSE, 
            message = TRUE, 
            cache = FALSE, 
            dpi = 300 
            )

# Function for numbers in rmarkdown
comma <- function(x) format(x, digits = 2, big.mark = ",")
```

Results file uploaded: __`r basename(rfile)`__. The data contains `r nrow(results)` rows and `r ncol(results)` columns.

Checks passed:

* All column names matched template for both files
* Values in the required columns were present
* Character length of values in columns were not too large
* P-values are between 0 and 1
* Values in CpG column of the results file look like CpGs
* Standard errors are positive (if present)

Here is a look at the top 10 strongest associations in your results (ordered by p-value)

```{r results-check, results="asis"}
top_hits <- results[order(results$P), ]
top_hits <- head(top_hits, n = 10)
top_hits <- apply(top_hits, 2, comma)
knitr::kable(top_hits, digits = Inf)
```