# Detailed EWAS report
### `r Sys.Date()`

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
pkgs <- c("knitr")
lapply(pkgs, require, character.only = TRUE)

knitr::opts_chunk$set(echo = FALSE, 
            warning = FALSE, 
            message = TRUE, 
            cache = FALSE, 
            dpi = 300 
            )

# Function for numbers in rmarkdown
comma <- function(x) format(x, digits = 2, big.mark = ",")
```

## Checklist

1. Range of summary statistics are reasonable, e.g. standard errors aren't all above 1000.
2. Number of associations are reasonable.
3. Check the EFO term(s). If it's not present this should be filled in.
4. Array entered and number of rows in results matches.
5. If trait has well established cpg-trait associations, they are present in top hits (e.g. cg05575921 is present for smoking EWAS).
6. If you haven't done so already, have check through the studies table that would have been sent back to the user in an email after they originally uploaded data. 

## Summary of results

* P-value range: `r paste(comma(range(full_results$P)), collapse = ", ")`
* N sites with P<1e-7: `r n_p7`
* N sites with P<1e-4: `r n_p4`
* SE range is `r range(full_results$SE)`
* Number of effect estimate outliers: `r n_beta_outliers`
* EFO term entered for `r studies$Trait` is: `r studies$EFO`
* Array: `r studies$Methylation_Array`
* Rows in results: `r nrow(results)` 

## P-values

Here is a look at the top 10 strongest associations in your results (ordered by p-value):

```{r results-check, results="asis"}
knitr::kable(top_hits, digits = Inf)
```

```{r man-plot, results="asis"}
manhattan.plot
```

```{r qq-plot, results="asis"}
qq.plot
```

## Beta + SE distributions

```{r beta-histogram, results="asis"}
if (betas_present) hist(full_results$Beta)
```

```{r se-histogram, results="asis"}
if (se_present) hist(full_results$SE)
```
